---
title: "Stat_analysis_zebra_SR"
author: "Justine Boutry"
date: "2023-01-22"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE, warning = F)
library(readr)
library(GGally)
library(glmmTMB)
library(sjPlot)
library(DHARMa)
library(MuMIn)
library(MASS)
library(lme4)
```

# Determination of sex in embryos injected with an oncogenic plasmid and in their offspring

```{r cars}
F0 <- read_delim("F0.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

F0$SR <- F0$Nb_Female/F0$Nb_tot

F0$Aquarium <- as.factor(F0$Aquarium)
F0$GFP <- as.factor(F0$GFP)

summary(F0)
```

# Analysis of the sexe ratios in the F0

## Graphical representation

```{r}
# Confident interval calculations

prop.test(sum(F0$Nb_Female[F0$GFP=="Pos"]), sum(F0$Nb_tot[F0$GFP=="Pos"]))
prop.test(sum(F0$Nb_Female[F0$GFP=="Neg"]), sum(F0$Nb_tot[F0$GFP=="Neg"]))

df <- data.frame(GFP=c("Neg", "Pos"),
                moy=c(
                  mean(F0[F0$GFP=="Neg",]$Nb_Female/F0[F0$GFP=="Neg",]$Nb_tot),
                  mean(F0[F0$GFP=="Pos",]$Nb_Female/F0[F0$GFP=="Pos",]$Nb_tot)),
                born.inf=c(0.4077056,0.4338454),
                born.sup=c(0.5405642,0.5988334))

(a <- ggplot(data = df, aes(y=moy, x=GFP, fill=GFP))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=born.inf, ymax=born.sup), width=.2,
                 position=position_dodge(.9))+
  #geom_point(aes(color=GFP), size = 3, alpha = .8)+
  geom_hline(yintercept=0.5, lty=2, color="darkgrey")+
  ylim(0,0.8)+
  labs(color="Expression of the \noncogenic plasmid", fill="Expression of the \noncogenic plasmid")+
  ylab("Proportion of females\n")+
  xlab("\nF0 generation\nfrom injected emrbyos")+
  scale_fill_manual(values = c("lightgrey", "darkred"))+
  scale_color_manual(values = c("lightgrey", "darkred"))+
  theme_minimal()+
  theme(
    panel.grid.major.x = element_blank(),
        axis.title.x=element_text(vjust=0,  
                                  size=20),  # X axis title
        axis.title.y=element_text(size=20, angle = 90),  # Y axis title
        axis.text.x=element_text(size=15,
                                 vjust=.5),  # X axis text
        axis.text.y=element_text(size=12))+
   ggsignif::geom_signif(comparisons = list(c("Neg", "Pos")), map_signif_level = TRUE, y_position = 0.7))
 
ggsave("SR_F0.jpg", a, device = "jpg")
```

### Random effect selection

```{r, echo=T, message=FALSE}
mod_SR_RN <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~GFP*Nb_tot+ (1|Exp/Aquarium), family = "binomial", REML = F)
mod_SR_R1 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~GFP*Nb_tot+ (1|Aquarium), family = "binomial", REML = F)
mod_SR_R0 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~GFP*Nb_tot, family = "binomial", REML = F)
AICc(mod_SR_RN, mod_SR_R1,  mod_SR_R0)
Weights(AICc(mod_SR_RN, mod_SR_R1,  mod_SR_R0))
```

### Fixed effects selection

```{r, echo=T, message=FALSE}
mod_SR_1 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~GFP*Nb_tot, family = "binomial", REML = T)
mod_SR_2 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~GFP+Nb_tot, family = "binomial", REML = T)
mod_SR_3 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~GFP, family = "binomial", REML = T)
mod_SR_4 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~Nb_tot, family = "binomial", REML = T)
mod_SR_0 <- glmmTMB(data=F0, cbind(Nb_Female, Nb_male)~1, family = "binomial", REML = T)

AICc(mod_SR_0, mod_SR_2,  mod_SR_3, mod_SR_4, mod_SR_1)
Weights(AICc(mod_SR_0, mod_SR_2,  mod_SR_3, mod_SR_4, mod_SR_1))

simulateResiduals(mod_SR_0, plot=T)
```

None of the effects measured in the study explains the sex ratio of F0 individuals from eggs injected with plasmid or not.

### Power analysis

```{r, echo=FALSE}

(muSRCtrl <- mean(F0$SR[F0$GFP=="Neg"]))
(nbCtrl <- length(F0$SR[F0$GFP=="Neg"]))
(nCtrl <-sum(F0$Nb_tot[F0$GFP=="Neg"]))

(muSRTum <- mean(F0$SR[F0$GFP=="Pos"]))
(nbTum <- mean(length(F0$SR[F0$GFP=="Pos"])))
(nTum <- sum(F0$Nb_tot[F0$GFP=="Pos"]))

i=0
q1 <- NULL
q3 <- NULL
Biais <- NULL

Results <- matrix(0, 1000, 40)
Ratios_accept <- NULL

# Avec un t - test
i=1
for (j in 1 : 1000){
  for (i in 1:40){
  FemCtrl <- rbinom(n=1, size= nCtrl, p= muSRCtrl)
  FemTum  <- rbinom(n=1, size= nTum, p= (muSRCtrl+(i*0.01)))
  Results[j, i] <- prop.test(c(FemCtrl, FemTum), n=c(nCtrl, nTum))$p.value
  q1[i] <- quantile(Results[,i], 0.025)
  q3[i] <- quantile(Results[,i], 0.975)
  }
}

pm <- apply(Results, 2, mean)
biais <- seq(1*0.01:40*0.01)

plot(biais, pm, type = "l", col = "red", 
            main="Robustness analysis",
            xlab="Sex Ratio Biais (%)",
            ylab = "Average p-value (1 000 simulations)",
          ylim=c(0,0.30),
          xlim=c(0,40))+abline(h=0.05, lty=3, col="grey")


```

# Impact of the cancer development on the offspring F1

```{r}
F1 <- read_delim("F1.txt", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE)

F1$Rep <- as.factor(F1$Rep)
F1$Breeding_date <- as.factor(F1$Breeding_date)
F1$GFP <- as.factor(F1$GFP)
F1$parent <- as.factor(F1$parent)
F1$SR <- F1$femelles/F1$Total

summary(F1)
```

## Size of the pond 3 days post fecundation

```{r}
(a <- ggplot(data = F1, aes(y= F1$Nb_eclo_3d, x=GFP, fill=GFP))+
  #geom_hline(yintercept=0.5, lty=2, color="darkgrey")+
  geom_boxplot(alpha=.7)+
  geom_point(aes(color=GFP), size = 3)+
  #ylim(0.1,0.9)+
  labs(color="Expression of the \noncogenic plasmid", fill="Expression of the \noncogenic plasmid")+
  ylab("Number of viable embryos (3dpf)\n")+
  xlab("\nF1 generation \n from tumorous mothers")+
  scale_fill_manual(values = c("lightgrey", "darkred"))+
  scale_color_manual(values = c("lightgrey", "darkred"))+
  theme_minimal()+
  theme(
    panel.grid.major.x = element_blank(),
        axis.title.x=element_text(vjust=0,  
                                  size=20),  # X axis title
        axis.title.y=element_text(size=20, angle = 90),  # Y axis title
        axis.text.x=element_text(size=15,
                                 vjust=.5),  # X axis text
        axis.text.y=element_text(size=12))+ 
   ggsignif::geom_signif(comparisons = list(c("neg", "pos")), map_signif_level = TRUE, y_position = 320)
 )
ggsave("SR_F1_Nb_egg.jpg", a, device = "jpg")
```

### Random effects selection

```{r, echo=FALSE, message=FALSE}
mod_3dp_RN <- glmmTMB(data=F1, Nb_eclo_3d~GFP+ (1|Rep/Breeding_date), family = "poisson", REML = F)
mod_3dp_R1 <- glmmTMB(data=F1, Nb_eclo_3d~GFP+ (1|Breeding_date), family = "poisson", REML = F)
mod_3dp_R0 <- glmmTMB(data=F1, Nb_eclo_3d~GFP, family = "poisson", REML = F)
AICc(mod_3dp_RN, mod_3dp_R1,  mod_3dp_R0)
Weights(AICc(mod_3dp_RN, mod_3dp_R1,  mod_3dp_R0))
```

Here it is important to take into account the influence of the breeding date on the total number of eggs released.

### Fixed effects selection

```{r, echo=FALSE, message=FALSE}
mod_3dp_R1 <- glmmTMB(data=F1, Nb_eclo_3d~GFP+ (1|Breeding_date), family = "poisson", REML = T)
mod_3dp_R1_0 <- glmmTMB(data=F1, Nb_eclo_3d~1+ (1|Breeding_date), family = "poisson", REML = T)

AICc(mod_3dp_R1, mod_3dp_R1_0)
Weights(AICc(mod_3dp_R1, mod_3dp_R1_0))

simulateResiduals(mod_3dp_R1, plot=T)

mod_3dp_R1 <- glmmTMB(data=F1, Nb_eclo_3d~GFP+ (1|Breeding_date), family = "poisson", REML = F)
tab_model(mod_3dp_R1, show.intercept = F)
```

The cancerous females lays more eggs than the non cancerous ones.

## Total number of individuals surviving 13 weeks

### Graphical representation

```{r, echo=FALSE, message=FALSE}

(a <- ggplot(data = F1, aes(y= F1$Total, x=GFP, fill=GFP))+
  #geom_hline(yintercept=0.5, lty=2, color="darkgrey")+
  geom_boxplot(alpha=0.7)+
  geom_point(aes(color=GFP), size = 3)+
  #ylim(0.1,0.9)+
  labs(color="Expression of the \noncogenic plasmid", fill="Expression of the \noncogenic plasmid")+
  ylab("Average number of descendant (13th week)")+
  xlab("\nF1 generation \n from tumorous mothers")+
  scale_fill_manual(values = c("lightgrey", "darkred"))+
  scale_color_manual(values = c("lightgrey", "darkred"))+
  theme_minimal()+
  theme(
    panel.grid.major.x = element_blank(),
        axis.title.x=element_text(vjust=0,  
                                  size=20),  # X axis title
        axis.title.y=element_text(size=20, angle = 90),  # Y axis title
        axis.text.x=element_text(size=15,
                                 vjust=.5),  # X axis text
        axis.text.y=element_text(size=12))+ 
   ggsignif::geom_signif(comparisons = list(c("neg", "pos")), map_signif_level = TRUE, y_position = 110)
 )
ggsave("SR_F1_nbTot.jpg", a, device = "jpg")
```

### Random effects selection

```{r, echo=FALSE, message=FALSE}
mod_Mat_R1 <- glmmTMB(data=F1, cbind(Total, Nb_eclo_3d)~GFP+ (1|Rep/Breeding_date), family = "binomial", REML = F)
mod_Mat_R2 <- glmmTMB(data=F1, cbind(Total, Nb_eclo_3d)~GFP+ (1|Breeding_date), family = "binomial", REML = F)
mod_Mat_R0 <- glmmTMB(data=F1, cbind(Total, Nb_eclo_3d)~GFP, family = "binomial", REML = F)
AICc(mod_Mat_R1, mod_Mat_R2,  mod_Mat_R0)
Weights(AICc(mod_3dp_RN, mod_3dp_R1,  mod_3dp_R0))
```

Here it is important to take into account the impact of the breeding date, nested into the experimental trial.

### Fixed effects selection

```{r, echo=FALSE, message=FALSE}
mod_Mat_R1 <- glmmTMB(data=F1, cbind(Total, Nb_eclo_3d)~GFP+ (1|Rep/Breeding_date), family = "binomial", REML = T)
mod_Mat_R1_0 <- glmmTMB(data=F1, cbind(Total, Nb_eclo_3d)~1+ (1|Rep/Breeding_date), family = "binomial", REML = T)
AICc(mod_Mat_R1, mod_Mat_R1_0)
Weights(AICc(mod_Mat_R1, mod_Mat_R1_0))

mod_Mat_R1 <- glmmTMB(data=F1, cbind(Total, Nb_eclo_3d)~GFP+ (1|Rep/Breeding_date), family = "binomial", REML = F)
tab_model(mod_Mat_R1, show.intercept = F)
```

The cancerous females have an increased ratio of eggs layed reaching the adult age compared to the healthy ones.

## Sexe-ratio in the F1 

Crosses were made every day. Each female was presented to a male several times until mating occurred. The eggs were then collected and the offspring sexed at 13 weeks.

### Random effects selection

```{r, echo=FALSE, message=FALSE}
mod_SRF1_R1 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Nb_eclo_3d+ (1|Rep/Breeding_date), family = "binomial", REML = F)
mod_SRF1_R2 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = F)
mod_SRF1_R0 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Nb_eclo_3d, family = "binomial", REML = F)

AICc(mod_SRF1_R1,mod_SRF1_R2,mod_SRF1_R0)
Weights(AICc(mod_SRF1_R1,mod_SRF1_R2,mod_SRF1_R0))
```

There is non need to take into account random effects in this analysis.

### Fixed effects selection

```{r, echo=TRUE, message=FALSE}
mod_SR_1 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Total*Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)

mod_SR_2 <- glmmTMB(data=F1, cbind(femelles, males)~GFP+Total*Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_3 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Total+Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_4 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Nb_eclo_3d+Total+ (1|Breeding_date), family = "binomial", REML = T)

mod_SR_5 <- glmmTMB(data=F1, cbind(femelles, males)~Total*Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_6 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Total+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_7 <- glmmTMB(data=F1, cbind(femelles, males)~GFP*Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)

mod_SR_8 <- glmmTMB(data=F1, cbind(femelles, males)~Total+Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_9 <- glmmTMB(data=F1, cbind(femelles, males)~GFP+Total+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_10 <- glmmTMB(data=F1, cbind(femelles, males)~GFP+Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_11 <- glmmTMB(data=F1, cbind(femelles, males)~GFP+Total+Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)

mod_SR_12 <- glmmTMB(data=F1, cbind(femelles, males)~Nb_eclo_3d+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_13 <- glmmTMB(data=F1, cbind(femelles, males)~GFP+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_14 <- glmmTMB(data=F1, cbind(femelles, males)~Total+ (1|Breeding_date), family = "binomial", REML = T)
mod_SR_15 <- glmmTMB(data=F1, cbind(femelles, males)~1+ (1|Breeding_date), family = "binomial", REML = T)

AICc(mod_SR_1, mod_SR_2, mod_SR_3, mod_SR_4,
     mod_SR_5, mod_SR_6, mod_SR_7, mod_SR_8, mod_SR_9,
     mod_SR_10, mod_SR_11, mod_SR_12, mod_SR_13, mod_SR_14,
     mod_SR_15)
Weights(AICc(mod_SR_1, mod_SR_2, mod_SR_3, mod_SR_4,
     mod_SR_5, mod_SR_6, mod_SR_7, mod_SR_8, mod_SR_9,
     mod_SR_10, mod_SR_11, mod_SR_12, mod_SR_13, mod_SR_14,
     mod_SR_15))

mod_SR_15 <- glmmTMB(data=F1, cbind(femelles, males)~1+ (1|Breeding_date), family = "binomial", REML = F)

simulateResiduals(mod_SR_15, plot = T)
```

There is absolutely no variations of the sex ratio than we can attributed to any effects we measured.

```{r}

(a <- ggplot(data = F1, aes(y= femelles/Total, x=GFP, fill=GFP))+
  geom_hline(yintercept=0.5, lty=2, color="darkgrey")+
  geom_boxplot(alpha=0.7)+
  geom_point(aes(color=GFP), size = 3, alpha = .8)+
  ylim(0.1,0.9)+
  labs(color="Expression of the \noncogenic plasmid", fill="Expression of the \noncogenic plasmid")+
  ylab("Proportion of females\n")+
  xlab("\nF1 generation\n from tumorous females")+
  scale_fill_manual(values = c("lightgrey", "darkred"))+
  scale_color_manual(values = c("lightgrey", "darkred"))+
  theme_minimal()+
  theme(
    panel.grid.major.x = element_blank(),
        axis.title.x=element_text(vjust=0,  
                                  size=20),  # X axis title
        axis.title.y=element_text(size=20, angle = 90),  # Y axis title
        axis.text.x=element_text(size=15,
                                 vjust=.5),  # X axis text
        axis.text.y=element_text(size=12))+ 
   ggsignif::geom_signif(comparisons = list(c("neg", "pos")), map_signif_level = TRUE, y_position = 0.8)
 )
ggsave("SR_F1T.jpg", a, device = "jpg")
```

### Power analysis

```{r}

muSRCtrl <- mean(F1$SR[F1$GFP=="neg"])
muSRTum <- mean(F1$SR[F1$GFP=="pos"])

q1 <- NULL
q1.1 <- NULL
q3 <- NULL
q3.1 <- NULL
Biais <- NULL

Results <- matrix(0, 1000, 40)
Ratios_accept <- NULL

# Avec glm
i=1

for (j in 1 : 1000){
  for (i in 1:40){
  FemCtrl <- rbinom(n=20, size= 37, p= muSRCtrl)
  FemTum  <- rbinom(n=15, size= 53, p= (muSRCtrl+(i*0.01)))
  data <- data.frame(GFP=c(rep("Ctrl", 20), rep("Tum", 15)), 
                     Total=c(rep(37, 20), rep(53, 15)), 
                     Fem=c(FemCtrl, FemTum))
  Results[j, i] <- summary(glm(data = data, cbind(Fem, Total-Fem) ~ GFP, family = binomial))$coefficients[2,4]
  q1.1[i] <- quantile(Results[,i], 0.025)
  q3.1[i] <- quantile(Results[,i], 0.975)
  }
}

pm1 <- apply(Results, 2, mean)

# Avec un t - test

for (j in 1 : 1000){
  for (i in 1:40){
  FemCtrl <- rbinom(n=20, size= 37, p= muSRCtrl)
  FemTum  <- rbinom(n=15, size= 53, p= (muSRCtrl+(i*0.01)))
  data <- data.frame(GFP=c(rep("Ctrl", 20), rep("Tum", 15)), 
                     Total=c(rep(37, 20), rep(53, 15)), 
                     Fem=c(FemCtrl, FemTum))
  Results[j, i] <- t.test(FemCtrl/37, FemTum/53)$p.value
  q1[i] <- quantile(Results[,i], 0.025)
  q3[i] <- quantile(Results[,i], 0.975)
  }
}

pm <- apply(Results, 2, mean)
biais <- seq(1:40)

plot(biais, pm, type = "l", col = "red", 
            main="Analyse de robustesse",
            xlab="Biais de sexe ratio",
            ylab = "P.value",
          ylim=c(0,0.10))+
    points(biais, q1, type = "l", lty=2, col="red")+
    points(biais, q3, type = "l", lty=2, col="red")+
  points(biais, pm1, type="l",col="blue")+
  points(biais, q1.1, type="l", lty=2, col="blue")+
  points(biais, q3.1, type="l",lty=2, col="blue")+
  abline(h=0.05, lty=3, col="grey")

```

A sex ratio bias of around 10% could have been detected, and the effect was only due to the first experiment. If I only analyze this data for experiment number 2, there is no effect.

### Citations

```{r}
citation("readr")
citation("GGally")
citation("glmmTMB")
citation("sjPlot")
citation("DHARMa")
citation("MuMIn")
citation("MASS")

sessionInfo()
```


